---
title: "p8105_hw3_amh2325"
author: "Amanda Howarth"
date: "10/9/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(p8105.datasets)
library(dplyr)
library(viridis)
library(ggridges) 
library(patchwork)
```

## Loading instacart data 
```{r}
library(p8105.datasets)
data("instacart")
```
## Exploring Instacart dataset 

```{r}
instacart %>%
  count(aisle_id)%>%
  arrange(desc(n))
```
#There are 134 aisles. The most items are ordered from Aisle 83 (150609 items), Aisle 24 (150473 items), and 123 (78493 items).

#Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.
```{r}
instacart%>%
  count(aisle_id)%>%
  filter(n > 10000)%>%
  ggplot(aes(x = aisle_id, y = n)) + geom_point() +
  labs(title = "Number of Items Ordered in Each Grocery Aisle", 
                x = "Aisle ID",
                y = "Number of Items Ordered in the Aisle")
```
## Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.

```{r}
popular_items_table=
instacart %>%
  group_by(aisle) %>%
  count(product_name, name = "count") %>%
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>%
  mutate(rank = min_rank(desc(count))) %>%
  filter(min_rank(rank) < 4)
```

```{r}
apples_icecream_table =
  instacart %>%
  select(product_name, order_dow, order_hour_of_day) %>% 
  group_by(product_name, order_dow) %>%
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>% 
  summarise(mean_hour_of_day = mean(order_hour_of_day)) %>%
  mutate(order_dow = recode(order_dow, "0" = "sunday", "1" = "monday", "2" = "tuesday", "3" = "wednesday", "4" = "thursday", "5" = "friday", "6" = "saturday")) %>% 
  pivot_wider(names_from = "order_dow", values_from = "mean_hour_of_day")
```

## Problem 2
##Loading BRFSS data 
```{r}
data("brfss_smart2010")
```
#Cleaning BRFSS data 
```{r}
brfss_overall_clean = brfss_smart2010 %>%
  janitor::clean_names() %>%
  filter(topic == "Overall Health") %>%
  mutate_all(tolower) %>%
  filter(response %in% c("poor", "fair", "good", "very good", "excellent")) %>%
  mutate(response = forcats::fct_relevel(response, c("poor", "fair", "good", "very good", "excellent"))) %>%
  rename("state" = "locationabbr", "county" = "locationdesc")
```
## In 2002, which states were observed at 7 or more locations? What about in 2010?
```{r}
brfss_2002 = brfss_overall_clean %>%
  group_by(year, state) %>%
  summarize(n = n_distinct(geo_location)) %>%
  filter(year == "2002", n >= 7)

brfss_2010 = brfss_overall_clean %>%
  group_by(year, state) %>%
  summarize(n = n_distinct(geo_location)) %>%
  filter(year == "2010", n >= 7)
```
## In 2002, six states (CT[7], FL[7], MA[8], NC[7], NJ[8], PA[10]) were observed at more than seven locations. In 2010, fourteen states (CA[12], CO[7], FL[41], MA[9], MD[12], NC[12], NE[10], NJ[19], NY[9], OH[8], PA[7], SC[7], TX[16], WA[10]) were observed at more than seven locations. 

## Construct a dataset that is limited to Excellent responses, and contains, year, state, and a variable that averages the data_value across locations within a state. 
```{r}
brfss_excellent = brfss_overall_clean %>%
  mutate(data_value = as.numeric(data_value)) %>%
  filter(response %in% c("excellent")) %>% 
  group_by(state, year, response) %>%
    summarize(data_value_average = mean(data_value)) %>%
  select(year, state, data_value_average)
```
## Make a “spaghetti” plot of this average value over time within a state (that is, make a plot showing a line for each state across years – the geom_line geometry and group aesthetic will help).
```{r}
ggplot(brfss_excellent, aes(x = year, y = data_value_average, group = state, color = state)) + geom_point() + geom_line() 
  + labs(title = "Average Data Values Across Locations within a State from 2002 - 2010", 
           x = "Year",
           y = "Average data value")
```

## Make a two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State.
```{r}
brfss_plot2006_2010 = brfss_overall_clean %>%
  mutate(data_value = as.numeric(data_value)) %>%
  filter(state == "ny", year == "2006" | year == "2010") %>%
  select(state, year, response, data_value, geo_location)
``` 

```{r}
plot_2006 = 
  brfss_plot2006_2010 %>%
  filter(year == "2010") %>%
  ggplot(aes(x = year, y = data_value_average), group = geo_location, color = blue) + geo_point()
  
```
## Problem #3: Loading Accel_data. Tidying and Wrangling.
```{r}
accel_data = 
  read_csv(file = "./data/accel_data.csv")
```
```{r}
accel_data_clean = accel_data %>%
  janitor::clean_names() %>%
  mutate(day = str_to_lower(day)) %>%
    mutate(weekday_vs_weekend = case_when(
    day == "monday" ~ "weekday",
    day == "tuesday" ~ "weekday",
    day == "wednesday" ~ "weekday",
    day == "thursday" ~ "weekday",
    day == "friday" ~ "weekday",
    day == "saturday" ~ "weekend",
    day == "sunday" ~ "weekend")) %>% 
  mutate_if(is.numeric, round, digits = 2) %>% 
  rename(`day_of_week` = day) %>%
select(week, day_id, day_of_week, weekday_vs_weekend, everything())
```  
There are `r nrow(accel_data_clean)` observations and `r ncol(accel_data_clean)` variables in `accel_data_clean`. There is a variable "week" for which week the man wore the accelerometer, variable "day_id" indicating which day of the 35 days the man wore the accelerometer, variable "day" indicating which day of the week (Monday, Tuesday, etc.), and "weekday_vs_weekend" indicating if the day was during the week or on the weekend. Lastly, there are 1440 variables for activity_*, the activity counts for each minute of a 24-hour day starting at midnight.

## Aggregating accross minutes to create a total activity variable for each day, and creating a table showing these totals. 
```{r}
accel_total_table = accel_data_clean %>%
  pivot_longer(activity_1:activity_1440, names_to = "minute", values_to = "activity_total") %>%
  group_by(week, day_id, day_of_week) %>% 
  summarise(activity_total_day = sum(activity_total))
```
## Are any trends apparent? 
There are no trends apparent in the table. The variable "activity_total_day" increases and decreases with no trend or pattern.

## Making a single-panel plot that shows the 24-hour activity time courses for each day and using color to indicate day of the week. 
```{r}
accel_total_table %>%
  ggplot(aes(x = day_id, y = activity_total_day, color = day_of_week)) + geom_line() + labs(title = "Twenty four-hour activity time courses for each day ", 
           x = "Day",
           y = "Activity total")

```
## Are there any patterns or conclusions you can make based on this graph? 
From the graph above, I can conclude that the the man's physical activity peaked on a Monday, day 16, in the third week. The man's activity was at his lowest for two Saturdays in week 4 and week 5. For weekends, the man's physical activity at baseline (day 1) was greater than his activity at the end (day 35). For weekdays, the man's physical activity was reported to be higher than at baseline. 

