---
title: "p8105_hw3_amh2325"
author: "Amanda Howarth"
date: "10/9/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(p8105.datasets)
library(dplyr)
library(viridis)
library(ggridges) 
library(patchwork)
```

## Loading instacart data 
```{r}
library(p8105.datasets)
data("instacart")
```
## Exploring Instacart dataset 

```{r}
instacart %>%
  count(aisle_id)%>%
  arrange(desc(n))
```
#There are 134 aisles. The most items are ordered from Aisle 83 (150609 items), Aisle 24 (150473 items), and 123 (78493 items).

#Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.
```{r}
instacart%>%
  count(aisle_id)%>%
  filter(n > 10000)%>%
  ggplot(aes(x = aisle_id, y = n)) + geom_point() +
  labs(title = "Number of Items Ordered in Each Grocery Aisle", 
                x = "Aisle ID",
                y = "Number of Items Ordered in the Aisle")
```
## Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.

```{r}
popular_items_table=
instacart %>%
  group_by(aisle) %>%
  count(product_name, name = "count") %>%
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>%
  mutate(rank = min_rank(desc(count))) %>%
  filter(min_rank(rank) < 4)
```

```{r}
apples_icecream_table =
  instacart %>%
  select(product_name, order_dow, order_hour_of_day)%>% 
  group_by(product_name, order_dow) %>%
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream"))%>% 
  summarise(mean_hour_of_day = mean(order_hour_of_day)) %>%
  mutate(order_dow = recode(order_dow, "0" = "sunday", "1" = "monday", "2" = "tuesday", "3" = "wednesday", "4" = "thursday", "5" = "friday", "6" = "saturday"))%>% 
  pivot_wider(names_from = "order_dow", values_from = "mean_hour_of_day")
```

##Loading BRFSS data 
```{r}
data("brfss_smart2010")
```
  
```{r}
brfss_smart2010 %>%
  janitor::clean_names() %>%
  filter(topic == "Overall Health") %>%
  filter(response %in% c("Poor", "Fair", "Good", "Very Good", "Excellent")) %>%
  mutate(response = forcats::fct_relevel(response, c("Poor", "Fair", "Good", "Very Good", "Excellent")))

```
  


